---
title: "Crimes"
author: "Isaac Aguilar"
date: "2024-03-18"
output: pdf_document
---

## Problem

Predict and, consequently, reduce the rate of violent crimes within our city.
1. Factors that influence the rate of Violent Crimes in Chicago.
2. Explore all possible correlations.
3. Particularly vulnerable periods.
4. Any Recommendations to reduce Violent Crimes


```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(skimr)
library(vroom)
library(lubridate)
library(DescTools)  # For Mode function
library(forcats) # uses as_factor (keeps levels the way they are)
library(fastDummies)
library(corrplot) 
library(progress)
library(glmnet)
```

```{r}
setwd("C:/Users/oia22/Box/College/Competition")

```

```{r, message = FALSE, warning = FALSE}
crimes <- vroom("Crimes.csv") 
weather <- vroom("weather.csv")
full_moon <- vroom("full_moon.csv")
holidays <- vroom("holidays.csv")
crimes_c <- vroom("crimes_cleaned.csv")
```


```{r}
# Make Date in Crime a date type
crimes$Datetime <- floor_date(mdy_hms(crimes$Date), unit = "hours")
crimes$Date <- as_date(crimes$Datetime)


# Merge all datasets into one by date'
master <- merge(crimes, weather, by.x = "Date", by.y = "datetime", all = TRUE)
full_moon_dates <- dmy(full_moon$FullMoonDates)
# Add full moons data
master <-  master %>%
  mutate(is_full_moon = if_else(Date %in% full_moon_dates, 1, 0))
# Add holidays data
master <-  master %>%
  mutate(is_holiday = if_else(Date %in% holidays$Date, 1, 0))

# Add hourly weather data
master <- merge(master, crimes_c[, c("DateTime", "hourly.temperature_2m","hourly.relative_humidity_2m","hourly.apparent_temperature", "hourly.rain", "hourly.snowfall", "hourly.snow_depth", "hourly.cloud_cover", "hourly.wind_speed_10m","hourly.wind_gusts_10m", "hourly.is_day","hourly.shortwave_radiation_instant", "hourly.direct_radiation_instant","Violent","Unemployment")],
                   by.x = "Datetime", by.y = "DateTime",
                   all.x = TRUE)
```

```{r}
# Filter from 2010 to end of 2023
master <- master %>% filter(Year >= 2010, Year < 2024)
```


```{r}
# Extract time variables of interest.
master$month <- month(master$Datetime, label=T)
master$week <- week(master$Datetime)
master$DayOfWeek <- wday(master$Datetime, label = TRUE)
master$hour <- hour(master$Datetime)
master$DayofMonth <- day(master$Datetime)
```


```{r}
# Identify violent crimes
names_v <- c("RAPE","MURDER","ASSAULT","CRIM SEXUAL ASSAULT	", "BATTERY", "CRIMINAL SEXUAL ASSAULT", "HOMICIDE", "HUMAN TRAFFICKING","INTIMIDATION", "KIDNAPPING", "ROBBERY", "SEX OFFENSE", "STALKING","OFFENSE INVOLVING CHILDREN") # Names of violent crimes
master$is_violent_crime <- ifelse(master$`Primary Type` %in% names_v | "VIOLENCE" %in% master$Description, 1, 0)

# Filter out non violent crimes
violent_crimes <- master %>% filter(is_violent_crime == 1) %>% distinct(`Primary Type`)
non_violent_crimes <- master %>% filter(is_violent_crime != 1) %>% distinct(`Primary Type`)
master <- master %>% filter(is_violent_crime == 1)
```

```{r}
# Remove unnecessary columns (Based on judgement)
rm_cols <- c("ID", "Case Number","Block","IUCR","Description","District","FBI Code","X Coordinate","Y Coordinate", "Updated On", "Location","name", "tempmax","tempmin","temp","feelslikemax" ,"feelslikemin","feelslike", "humidity", "precip","precipprob" ,"precipcover","preciptype","snow","snowdepth","windgust","windspeed","winddir","cloudcover","solarradiation","solarenergy","uvindex","severerisk","sunrise","sunset", "conditions","description","icon","stations", "Violent", "Date","is_violent_crime" )
master <- master[, !names(master) %in% rm_cols]
```

```{r}
# Rename some columns (weather data)
new_master <- master %>%
  rename("temperature" = hourly.temperature_2m,
         "humidity" = hourly.relative_humidity_2m,
         "feels_like" = hourly.apparent_temperature,
         "rain" = hourly.rain,
         "snowfall" = hourly.snowfall,
         "snow_depth" = hourly.snow_depth,
         "cloud_cover" = hourly.cloud_cover,
         "wind_speed" = hourly.wind_speed_10m,
         "wind_gusts" = hourly.wind_gusts_10m,
         "is_day" = hourly.is_day,
         "shortwave_radiation" = hourly.shortwave_radiation_instant,
         "direct_radiation" = hourly.direct_radiation_instant)
```



```{r}
# Order dataframe by Datetime and place columns in most logical order (Group them by time variable, spacial variables, etc)
new_master <- new_master %>%
  select(Datetime,Year,month,week,DayOfWeek,DayofMonth,hour,is_holiday,is_full_moon,moonphase,temperature,humidity,feels_like,rain,snowfall,snow_depth,cloud_cover,wind_speed,wind_gusts,is_day,shortwave_radiation,direct_radiation,dew,sealevelpressure,visibility,`Primary Type`,Arrest,Beat,Domestic,`Location Description`,Ward,`Community Area`,Latitude,Longitude,Unemployment) %>%
  arrange(Datetime)
```

**Questions before Grouping & Summarizing**

1. moonphase turn into factor?
```{r}
table(new_master$moonphase)
# Answer is no (Continuous variable)
```
2 is_day turn into factor?
```{r}
table(new_master$is_day)
# answer is yes
```

3. Turn into factors (Beat, Ward, Community Area)?
```{r}
table(new_master$Beat)
# Answer is yes
table(new_master$Ward)
# Answer is yes
table(new_master$`Community Area`)
# Answer is yes
```

4. Should we remove Latitude and Longitude?
```{r}
new_master <- new_master %>% filter(Longitude > -91) # Remove points that are not in the area (Longitude might be wrong)
new_master <- new_master %>% filter(Beat > 1100, Beat < 1200) # Remove rows with this info (no longitude and latitude recorded)
# Missing values for Community Area
new_master %>% filter(Latitude < 41.88, Latitude > 41.87, Longitude < -87.7, Longitude > -87.71) %>% select(`Community Area`) %>% distinct()
# After reviewing the map NAs from Community Area come from area #27
new_master <- new_master %>% mutate(`Community Area` = ifelse(is.na(`Community Area`), 27, `Community Area`))

# Ward, Community Area
filtered <- new_master %>% filter(Beat == 1111)
for (i in unique(new_master$Beat)) {
  filtered <- new_master %>% filter(Beat == i)
  
  # Create scatter plot for each Beat
  plot <- ggplot(filtered, aes(x = Longitude, y = Latitude)) +
    geom_point() +
    labs(x = "Longitude", y = "Latitude") +
    ggtitle(paste("Scatter Plot of Longitude vs Latitude for Beat", i)) +
    xlim(min(new_master$Longitude), max(new_master$Longitude)) +
    ylim(min(new_master$Latitude), max(new_master$Latitude))
  
  # Print each plot
  print(plot)
}

# Remove Latitude and Longitude for the grouped data, everything is captured in the other variables
new_master <- new_master %>% select(-Longitude, -Latitude, -`Location Description`, -Ward, -`Community Area`)
# Remove Datetime (we can create it back later with the columns we extracted from it)
new_master <- new_master %>% select(-Datetime)
```


```{r}
# Perform findings from questions
# new_master$is_day <- as_factor(new_master$is_day)
new_master$Beat <- as_factor(new_master$Beat)
new_master$Ward <- as_factor(new_master$Ward)
new_master$`Community Area` <- as_factor(new_master$`Community Area`)
new_master$`Primary Type` <- as_factor(new_master$`Primary Type`)
#new_master$is_holiday <- as_factor(new_master$is_holiday)
#new_master$is_full_moon <- as_factor(new_master$is_full_moon)
#new_master <- mutate_if(new_master, is.logical, as.factor)
new_master <- new_master %>% mutate(`Location Description` = ifelse(is.na(`Location Description`), 'UNKOWN', `Location Description`)) # Fix missing values from Location Description
new_master$`Location Description` <- as_factor(new_master$`Location Description`)
```

```{r}
# Reduce number of Location Description categories by grouping the ones with low frequency into an "ANOTHER" category

# If it contains CTA, just name it CTA
new_master <- new_master %>% mutate(`Location Description` = if_else(str_detect(`Location Description`, "CTA"), "CTA", `Location Description`))
new_master <- new_master %>% mutate(`Location Description` = if_else(str_detect(`Location Description`, "SCHOOL"), "SCHOOL, PUBLIC, BUILDING", `Location Description`))
new_master <- new_master %>% mutate(`Location Description` = if_else(str_detect(`Location Description`, "VEHICLE"), "VEHICLE", `Location Description`))
new_master <- new_master %>% mutate(`Location Description` = if_else(str_detect(`Location Description`, "AUTO"), "VEHICLE", `Location Description`))
category_percentages <- new_master %>%
  group_by(`Location Description`) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(percentage))

threshold_percentage <- .04

# Get the categories with percentages below the threshold
categories_to_group <- category_percentages %>%
  filter(percentage < threshold_percentage) %>%
  pull(`Location Description`)

new_master <- new_master %>%
  mutate(`Location Description` = ifelse(`Location Description` %in% categories_to_group, "OTHER", `Location Description`))
```


```{r}
# Make factor variables with multiple variables 
long_master <- dummy_cols(new_master, select_columns = c("Primary Type", "Beat", "Ward", "Community Area", "Location Description"))
long_master <- long_master %>% select(-c("Primary Type", "Beat", "Ward", "Community Area", "Location Description"))
```



```{r}
# Group and summarize to get # of crimes per hour of the year.
get_mean <- c("moonphase", "temperature", "humidity", "feels_like", "rain", 
                         "snowfall", "snow_depth", "cloud_cover", "wind_speed", 
                         "wind_gusts", "shortwave_radiation", "direct_radiation", 
                         "dew", "sealevelpressure", "visibility", "Unemployment")

final <- long_master %>%
  group_by(Year, month, week, DayOfWeek, DayofMonth, hour) %>%
  summarize(across(everything(),
                   ~ if_else(any(names(.) %in% get_mean),
                             mean(., na.rm = TRUE), 
                             sum(., na.rm = TRUE))),
             n_crimes = n())

# Save the final dataframe
write.csv(final, "final_table.csv", row.names = FALSE)
```
```{r}
# Load the saved object back into R
final_raw <- read_csv(file = "final_table.csv")
final_raw$hour <- as_factor(final_raw$hour)
final <- final_raw %>% select(!(matches("Location Description.*|Ward.*|Community Area.*|Primary Type.*|month.*"))) %>%
  select(-c("is_full_moon", "moonphase", "temperature", "humidity",
                        "feels_like", "rain", "snowfall", "cloud_cover",
                        "wind_gusts", "shortwave_radiation", "direct_radiation",
                        "dew", "visibility", "Arrest", "Domestic", "week"))
```
```{r}

```


## Select Features to use

```{r}
# Define exploratory and predictive
y <- final$n_crimes
X <- model.matrix(n_crimes~.+DayOfWeek*hour, data = final)

# Model
lasso.pois <- cv.glmnet(X, y, family = "poisson", alpha = 1, verbose = TRUE)
lambda.star <- lasso.pois$lambda.min # Best lambda
coefficients <- coef(lasso.pois, s = lambda.star)
coefficients
```

```{r}
# Check for Autocorrelation
predicted <- predict(lasso.pois, newx = X, s = lambda.star, type = "response") # Predicted values
residuals <- y - predicted # Residuals
acf(residuals)
mean(residuals^2) # Compute Mean Squared Error (MSE)
# MSE without Location Description.*|Ward.*|Community Area.*|Primary Type.* ==> 44.81995
# MSE without Location Description.*|Ward.*|Community Area.*|Primary Type.*|month.* ==> 39.4265
# MSE without Location Description.*|Ward.*|Community Area.*|Primary Type.*|month.* & other weather data ==> 33.76191
```

Are factors different depending on Primary Type?
```{r}
vc_type <- unique(new_master$`Primary Type`)

# Create 11 models 
```


## Select Best Model for prediction

```{r}
# Split between Train and Test

```


```{r}
# Poisson Regression

```


```{r}
# Time Series

```

```{r}
# Random Forest

```


## Predictions csv for Dashboard

```{r}
# Select best model and do predictions for Jan 2024 to March 2024


```

```{r}
# Save csv
```

